---
title: "Effect_size_differences"
author: "Lauren Blake"
date: "November 27, 2017"
output: html_document
---

This is a script to evaluate the effect sizes of the 2 linear models. 

## Load in data 


```{r chunk-options, include=FALSE}
source("chunk-options.R")
```


```{r}
# import sample labels
samples <- read.delim("../data/Sample_info_RNAseq_RIN.txt")
# expression
exprs <- read.table("../data/human_chimp_orth_exp_methyl_7725_hum.txt", sep="")
# methylation data
methyl <- read.csv("../data/chimp_human_orth_7725_avg_methyl_per_ts_gene.txt", sep="")

# Normalized gene expression data
cpm.voom.cyclic <- readRDS("../data/human_chimp_orth_cpm_voom_cyclic.rds")

# Load libraries
library("edgeR")
library("limma")
library("plyr")
library("ashr")
```

## Function that will perform the interspecies analysis- DE genes
```{r}

effect_size_hc_species_DE <- function(two_species, FDR_level, FSR_level){
## Determine DE genes (for tissue analysis)

# Factors that will go in the lm
chimp_human_heart <- two_species
index_sample <- chimp_human_heart
FDR_level <- FDR_level

# Get variables
species <- as.data.frame(samples[,2])
tissue <- as.data.frame(samples[,3])
RIN <- as.data.frame(samples[,4])

# Sort tissue samples
  tissue <- as.data.frame(samples[,2])
  tissue <- tissue[chimp_human_heart,]
  tissue_no_extra <- droplevels.factor(tissue)

# Run the linear model in limma
  design <- model.matrix(~ as.factor(tissue_no_extra) + RIN[chimp_human_heart,])
  fit_all <- lmFit(cpm.voom.cyclic[,chimp_human_heart], design)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]

# Find which DE genes have methylation values associated with it
    human_chimp_heart <-  rownames(exprs) %in% HvC_Heart_fit_all_5perc$genes
  human_chimp_heart <- as.data.frame(human_chimp_heart)
  counts_genes_in <- cbind(exprs, human_chimp_heart)
  counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart ==
"TRUE")
  counts_genes_in_cutoff <- counts_genes_in_cutoff[,1:79]
  expression_values_only <- counts_genes_in_cutoff[,2:32]
  methylation_values_only <- counts_genes_in_cutoff[,49:79]
  dim(expression_values_only)
  dim(methylation_values_only)
  
# Rename is that we can use Joyce's code
exprs_subset <- expression_values_only
methyl_subset <- methylation_values_only
  
# check agreement between methylation data and expression data
all.equal(rownames(methyl_subset), rownames(exprs_subset))
all.equal(colnames(methyl_subset), colnames(exprs_subset))

# <---- sample data matching
samples_sub <- samples[samples$Sample_ID %in% colnames(exprs_subset),]


exprs_pair <- exprs_subset[,index_sample]
methyl_pair <- methyl_subset[,index_sample]
species <- droplevels.factor(samples$Species[index_sample])
tissue <- droplevels.factor(samples$Tissue[index_sample])
RIN <- samples$RIN[index_sample]


# Extract parameters
N <- ncol(exprs_pair)
ngenes <- nrow(methyl_pair)
# <---- Model 1: exprs ~ tissue
# specify tissue coding
design_1 <- model.matrix(~species + RIN)
design_1[design_1[,2]==0,2] <- -1
model_1 <- lmFit(exprs_pair, design_1)

# <---- Model 3: exprs corrected for methylation ~ tissue
# specify design matrix
resid_exprs <- array(0, dim = dim(exprs_pair))
for (index in 1:nrow(exprs_pair)){
  resid_exprs[index,] <- lm(t(exprs_pair[index, ]) ~ t(methyl_pair[index, ]))$resid
}
rownames(resid_exprs) <- rownames(exprs_pair)
model_3 <- lmFit(resid_exprs, design_1)  


## Effect size differences

# get effect sizes
beta1 <- coef(model_1[,2])
beta3 <- coef(model_3[,2])
se_beta1 <- se_beta2 <- se_beta3 <- cov_beta13 <- vector("numeric", ngenes)
# <---- get variances of beta1^S tissue effect
se_beta1 <- model_1$sigma*sqrt((solve(t(design_1)%*%design_1))[2,2])
head(cbind(se_beta1,model_1$sigma*sqrt(model_1$cov.coefficients[2,2])))
# <---- get variances of beta2 methylation effect
for (g in 1:length(se_beta2)) {
  design_2g <- model.matrix(~ as.numeric(methyl_pair[g,]))  
  sigma_g <- model_1$sigma[g]
  se_beta2[index] <- sigma_g*sqrt((solve(t(design_2g)%*%design_2g))[2,2])
}
# <---- get variances of beta3 tissue effect
A <- solve(t(design_1)%*%design_1)%*%t(design_1)
contr.vector <- array(c(0,1), dim=c(2,1))
# compute beta3 by hand
for (g in 1:length(se_beta3)) {
  M_g <- t(methyl_pair[g,])
  design_2g <- model.matrix(~ as.numeric(M_g))  
  A_2g <- solve(t(design_2g)%*%design_2g)%*%t(design_2g)
  sigma_g <- model_1$sigma[g]
  var_beta2g <- (se_beta2^2)[g]
  var_part1 <- (se_beta1[g])^2
  var_part2 <- ( A%*%M_g%*%var_beta2g%*%t(M_g)%*%t(A) )[2,2]
  var_part3 <- ( 2*(sigma_g^2)*A%*%t(A_2g)%*%contr.vector%*%t(M_g)%*%t(A) )[2,2]
  se_beta3[g] <- sqrt(var_part1 + var_part2 + var_part3)
  }
# cov(beta1,beta3)
for (g in 1:length(cov_beta13)) {
  M_g <- t(methyl_pair[g,])
  design_2g <- model.matrix(~ as.numeric(M_g))  
  A_2g <- solve(t(design_2g)%*%design_2g)%*%t(design_2g)
  sigma_g <- model_1$sigma[g]
  var_part1 <- (se_beta1[g])^2
  cov_beta13[g] <- var_part1 - (sigma_g^2)*((A %*% t(A_2g) %*%contr.vector%*%t(M_g)%*%t(A))[2,2])
}
  
# cov(beta1-beta3)
cov_diff_sqrt <- sqrt(se_beta1^2 + se_beta3^2 - 2*cov_beta13)
beta_diff <- beta1-beta3

## Run adaptive shrinkage (with Matthew Stephens' package ashr)

fit_ash <- ash(as.vector(beta_diff), cov_diff_sqrt, mode=0)
names(fit_ash$result)

# lfsr is the minimum of PositiveProb and NegativeProb
# look at genes at various cut-off, say svalue < .01 etc. do they make sense?

ind_sig <- (fit_ash$result$svalue < FSR_level)
ind_sig[ind_sig == TRUE] <- "red"
ind_sig[ind_sig == FALSE] <- "black"

make_plot <- cbind(beta1-beta3, cov_diff_sqrt, ind_sig, fit_ash$result$lfdr)
colnames(make_plot) <- c("Diff", "SE", "Sign", "LFDR")
make_plot[,1] <- as.numeric(make_plot[,1])
make_plot[,2] <- as.numeric(make_plot[,2])
make_plot[,3] <- as.character(make_plot[,3])
make_plot[,4] <- as.numeric(make_plot[,4])

return(make_plot)
}
```

## Function that will perform the interspecies analysis- non-DE genes
```{r}

effect_size_hc_species_non_DE <- function(two_species, FDR_level, FSR_level){
## Determine DE genes (for tissue analysis)

# Factors that will go in the lm
chimp_human_heart <- two_species
index_sample <- chimp_human_heart
FDR_level <- FDR_level

# Get variables
species <- as.data.frame(samples[,2])
tissue <- as.data.frame(samples[,3])
RIN <- as.data.frame(samples[,4])

# Sort tissue samples
  tissue <- as.data.frame(samples[,2])
  tissue <- tissue[chimp_human_heart,]
  tissue_no_extra <- droplevels.factor(tissue)

# Run the linear model in limma
  design <- model.matrix(~ as.factor(tissue_no_extra) + RIN[chimp_human_heart,])
  fit_all <- lmFit(cpm.voom.cyclic[,chimp_human_heart], design)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]

# Find which DE genes have methylation values associated with it
    human_chimp_heart <-  rownames(exprs) %in% HvC_Heart_fit_all_5perc$genes
  human_chimp_heart <- as.data.frame(human_chimp_heart)
  counts_genes_in <- cbind(exprs, human_chimp_heart)
  counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart ==
"FALSE")
  counts_genes_in_cutoff <- counts_genes_in_cutoff[,1:79]
  expression_values_only <- counts_genes_in_cutoff[,2:32]
  methylation_values_only <- counts_genes_in_cutoff[,49:79]
  dim(expression_values_only)
  dim(methylation_values_only)
  
# Rename is that we can use Joyce's code
exprs_subset <- expression_values_only
methyl_subset <- methylation_values_only
  
# check agreement between methylation data and expression data
all.equal(rownames(methyl_subset), rownames(exprs_subset))
all.equal(colnames(methyl_subset), colnames(exprs_subset))

# <---- sample data matching
samples_sub <- samples[samples$Sample_ID %in% colnames(exprs_subset),]


exprs_pair <- exprs_subset[,index_sample]
methyl_pair <- methyl_subset[,index_sample]
species <- droplevels.factor(samples$Species[index_sample])
tissue <- droplevels.factor(samples$Tissue[index_sample])
RIN <- samples$RIN[index_sample]


# Extract parameters
N <- ncol(exprs_pair)
ngenes <- nrow(methyl_pair)
# <---- Model 1: exprs ~ tissue
# specify tissue coding
design_1 <- model.matrix(~species + RIN)
design_1[design_1[,2]==0,2] <- -1
model_1 <- lmFit(exprs_pair, design_1)

# <---- Model 3: exprs corrected for methylation ~ tissue
# specify design matrix
resid_exprs <- array(0, dim = dim(exprs_pair))
for (index in 1:nrow(exprs_pair)){
  resid_exprs[index,] <- lm(t(exprs_pair[index, ]) ~ t(methyl_pair[index, ]))$resid
}
rownames(resid_exprs) <- rownames(exprs_pair)
model_3 <- lmFit(resid_exprs, design_1)  


## Effect size differences

# get effect sizes
beta1 <- coef(model_1[,2])
beta3 <- coef(model_3[,2])
se_beta1 <- se_beta2 <- se_beta3 <- cov_beta13 <- vector("numeric", ngenes)
# <---- get variances of beta1^S tissue effect
se_beta1 <- model_1$sigma*sqrt((solve(t(design_1)%*%design_1))[2,2])
head(cbind(se_beta1,model_1$sigma*sqrt(model_1$cov.coefficients[2,2])))
# <---- get variances of beta2 methylation effect
for (g in 1:length(se_beta2)) {
  design_2g <- model.matrix(~ as.numeric(methyl_pair[g,]))  
  sigma_g <- model_1$sigma[g]
  se_beta2[index] <- sigma_g*sqrt((solve(t(design_2g)%*%design_2g))[2,2])
}
# <---- get variances of beta3 tissue effect
A <- solve(t(design_1)%*%design_1)%*%t(design_1)
contr.vector <- array(c(0,1), dim=c(2,1))
# compute beta3 by hand
for (g in 1:length(se_beta3)) {
  M_g <- t(methyl_pair[g,])
  design_2g <- model.matrix(~ as.numeric(M_g))  
  A_2g <- solve(t(design_2g)%*%design_2g)%*%t(design_2g)
  sigma_g <- model_1$sigma[g]
  var_beta2g <- (se_beta2^2)[g]
  var_part1 <- (se_beta1[g])^2
  var_part2 <- ( A%*%M_g%*%var_beta2g%*%t(M_g)%*%t(A) )[2,2]
  var_part3 <- ( 2*(sigma_g^2)*A%*%t(A_2g)%*%contr.vector%*%t(M_g)%*%t(A) )[2,2]
  se_beta3[g] <- sqrt(var_part1 + var_part2 + var_part3)
  }
# cov(beta1,beta3)
for (g in 1:length(cov_beta13)) {
  M_g <- t(methyl_pair[g,])
  design_2g <- model.matrix(~ as.numeric(M_g))  
  A_2g <- solve(t(design_2g)%*%design_2g)%*%t(design_2g)
  sigma_g <- model_1$sigma[g]
  var_part1 <- (se_beta1[g])^2
  cov_beta13[g] <- var_part1 - (sigma_g^2)*((A %*% t(A_2g) %*%contr.vector%*%t(M_g)%*%t(A))[2,2])
}
  
# cov(beta1-beta3)
cov_diff_sqrt <- sqrt(se_beta1^2 + se_beta3^2 - 2*cov_beta13)
beta_diff <- beta1-beta3

## Run adaptive shrinkage (with Matthew Stephens' package ashr)

fit_ash <- ash(as.vector(beta_diff), cov_diff_sqrt, mode=0)
names(fit_ash$result)

# lfsr is the minimum of PositiveProb and NegativeProb
# look at genes at various cut-off, say svalue < .01 etc. do they make sense?

ind_sig <- (fit_ash$result$svalue < FSR_level)
ind_sig[ind_sig == TRUE] <- "red"
ind_sig[ind_sig == FALSE] <- "black"

make_plot <- cbind(beta1-beta3, cov_diff_sqrt, ind_sig, fit_ash$result$lfdr)
colnames(make_plot) <- c("Diff", "SE", "Sign", "LFDR")
make_plot[,1] <- as.numeric(make_plot[,1])
make_plot[,2] <- as.numeric(make_plot[,2])
make_plot[,3] <- as.character(make_plot[,3])
make_plot[,4] <- as.numeric(make_plot[,4])

return(make_plot)
}
```

### Results: Human heart versus chimp heart

```{r}
######### Human heart versus chimp heart ###############

hc_heart_DE_make_plot <- effect_size_hc_species_DE(c(1, 5, 9, 13, 20, 24, 28), 0.05, 0.05)

plot(x=hc_heart_DE_make_plot[,1], y=hc_heart_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human versus chimp heart)",
     pch=16, cex=.6, col = hc_heart_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hc_heart_DE_make_plot[,3] == "red")
length(hc_heart_DE_make_plot[,3])

hc_heart_non_DE_make_plot <- effect_size_hc_species_non_DE(c(1, 5, 9, 13, 20, 24, 28), 0.05, 0.05)

plot(x=hc_heart_non_DE_make_plot[,1], y=hc_heart_non_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non DE genes in human versus chimp heart)",
     pch=16, cex=.6, col = hc_heart_non_DE_make_plot[,3])
legend("topright", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hc_heart_non_DE_make_plot[,3] == "red")
length(hc_heart_non_DE_make_plot[,3])


######### Human heart versus chimp heart ###############

hc_heart_DE_make_plot <- effect_size_hc_species_DE(c(1, 5, 9, 13, 20, 24, 28), 0.05, 0.01)

plot(x=hc_heart_DE_make_plot[,1], y=hc_heart_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human versus chimp heart)",
     pch=16, cex=.6, col = hc_heart_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.01", "s >= 0.01"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hc_heart_DE_make_plot[,3] == "red")
length(hc_heart_DE_make_plot[,3])

hc_heart_non_DE_make_plot <- effect_size_hc_species_non_DE(c(1, 5, 9, 13, 20, 24, 28), 0.05, 0.01)

plot(x=hc_heart_non_DE_make_plot[,1], y=hc_heart_non_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non DE genes in human versus chimp heart)",
     pch=16, cex=.6, col = hc_heart_non_DE_make_plot[,3])
legend("topright", legend=c("s < 0.01", "s >= 0.01"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hc_heart_non_DE_make_plot[,3] == "red")
length(hc_heart_non_DE_make_plot[,3])



######### Human heart versus chimp heart ###############

hc_heart_DE_make_plot <- effect_size_hc_species_DE(c(1, 5, 9, 13, 20, 24, 28), 0.1, 0.1)

plot(x=hc_heart_DE_make_plot[,1], y=hc_heart_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human versus chimp heart)",
     pch=16, cex=.6, col = hc_heart_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.1", "s >= 0.1"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hc_heart_DE_make_plot[,3] == "red")
length(hc_heart_DE_make_plot[,3])

hc_heart_non_DE_make_plot <- effect_size_hc_species_non_DE(c(1, 5, 9, 13, 20, 24, 28), 0.1, 0.1)

plot(x=hc_heart_non_DE_make_plot[,1], y=hc_heart_non_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non DE genes in human versus chimp heart)",
     pch=16, cex=.6, col = hc_heart_non_DE_make_plot[,3])
legend("topright", legend=c("s < 0.1", "s >= 0.1"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hc_heart_non_DE_make_plot[,3] == "red")
length(hc_heart_non_DE_make_plot[,3])

```

### Results: Human kidney versus chimp kidney

```{r}

hc_heart_DE_make_plot <- effect_size_hc_species_DE(c(2, 6, 10, 14, 17, 21, 25, 29), 0.05, 0.05)

plot(x=hc_heart_DE_make_plot[,1], y=hc_heart_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human versus chimp kidney)",
     pch=16, cex=.6, col = hc_heart_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

a <- sum(hc_heart_DE_make_plot[,3] == "red")
a
b <- length(hc_heart_DE_make_plot[,3])
b

hc_heart_non_DE_make_plot <- effect_size_hc_species_non_DE(c(2, 6, 10, 14, 17, 21, 25, 29), 0.05, 0.05)

plot(x=hc_heart_non_DE_make_plot[,1], y=hc_heart_non_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non DE genes in human versus chimp kidney)",
     pch=16, cex=.6, col = hc_heart_non_DE_make_plot[,3])
legend("topright", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

c <- sum(hc_heart_non_DE_make_plot[,3] == "red")
c
d <- length(hc_heart_non_DE_make_plot[,3])
d

test <- matrix(c(a,b,c,d), ncol = 2, byrow = FALSE)
prop.test(test, correct = TRUE)
```


### Results: Human liver versus chimp liver

```{r}
hc_heart_DE_make_plot <- effect_size_hc_species_DE(c(3, 7, 11, 15, 18, 22, 26, 30), 0.05, 0.05)

plot(x=hc_heart_DE_make_plot[,1], y=hc_heart_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human versus chimp liver)",
     pch=16, cex=.6, col = hc_heart_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hc_heart_DE_make_plot[,3] == "red")
length(hc_heart_DE_make_plot[,3])

hc_heart_non_DE_make_plot <- effect_size_hc_species_non_DE(c(3, 7, 11, 15, 18, 22, 26, 30), 0.05, 0.05)

plot(x=hc_heart_non_DE_make_plot[,1], y=hc_heart_non_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non DE genes in human versus chimp liver)",
     pch=16, cex=.6, col = hc_heart_non_DE_make_plot[,3])
legend("topright", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hc_heart_non_DE_make_plot[,3] == "red")
length(hc_heart_non_DE_make_plot[,3])

```

### Results: Human lung versus chimp lung

```{r}
######### Human heart versus chimp heart ###############

hc_heart_DE_make_plot <- effect_size_hc_species_DE(c(4, 8, 12, 16, 19, 23, 27, 31), 0.05, 0.05)

plot(x=hc_heart_DE_make_plot[,1], y=hc_heart_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human versus chimp lung)",
     pch=16, cex=.6, col = hc_heart_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hc_heart_DE_make_plot[,3] == "red")
length(hc_heart_DE_make_plot[,3])

hc_heart_non_DE_make_plot <- effect_size_hc_species_non_DE(c(4, 8, 12, 16, 19, 23, 27, 31), 0.05, 0.05)

plot(x=hc_heart_non_DE_make_plot[,1], y=hc_heart_non_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non DE genes in human versus chimp lung)",
     pch=16, cex=.6, col = hc_heart_non_DE_make_plot[,3])
legend("topright", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hc_heart_non_DE_make_plot[,3] == "red")
length(hc_heart_non_DE_make_plot[,3])

```



## Function that will perform the intertissue analysis- DE genes
```{r}

effect_size_hc_tissue_DE <- function(two_tissues, FDR_level, FSR_level){
## Determine DE genes (for tissue analysis)

# Factors that will go in the lm
chimp_human_heart <- two_tissues
index_sample <- chimp_human_heart
FDR_level <- FDR_level

# Get variables
species <- as.data.frame(samples[,2])
tissue <- as.data.frame(samples[,3])
RIN <- as.data.frame(samples[,4])

# Sort tissue samples
  tissue <- as.data.frame(samples[,3])
  tissue <- tissue[chimp_human_heart,]
  tissue_no_extra <- droplevels.factor(tissue)

# Run the linear model in limma
  design <- model.matrix(~ as.factor(tissue_no_extra) + RIN[chimp_human_heart,])
  fit_all <- lmFit(cpm.voom.cyclic[,chimp_human_heart], design)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]

# Find which DE genes have methylation values associated with it
    human_chimp_heart <-  rownames(exprs) %in% HvC_Heart_fit_all_5perc$genes
  human_chimp_heart <- as.data.frame(human_chimp_heart)
  counts_genes_in <- cbind(exprs, human_chimp_heart)
  counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart ==
"TRUE")
  counts_genes_in_cutoff <- counts_genes_in_cutoff[,1:79]
  expression_values_only <- counts_genes_in_cutoff[,2:32]
  methylation_values_only <- counts_genes_in_cutoff[,49:79]
  dim(expression_values_only)
  dim(methylation_values_only)
  
# Rename is that we can use Joyce's code
exprs_subset <- expression_values_only
methyl_subset <- methylation_values_only
  
# check agreement between methylation data and expression data
all.equal(rownames(methyl_subset), rownames(exprs_subset))
all.equal(colnames(methyl_subset), colnames(exprs_subset))

# <---- sample data matching
samples_sub <- samples[samples$Sample_ID %in% colnames(exprs_subset),]


exprs_pair <- exprs_subset[,index_sample]
methyl_pair <- methyl_subset[,index_sample]
species <- droplevels.factor(samples$Species[index_sample])
tissue <- droplevels.factor(samples$Tissue[index_sample])
RIN <- samples$RIN[index_sample]


# Extract parameters
N <- ncol(exprs_pair)
ngenes <- nrow(methyl_pair)
# <---- Model 1: exprs ~ tissue
# specify tissue coding
design_1 <- model.matrix(~tissue + RIN)
design_1[design_1[,2]==0,2] <- -1
model_1 <- lmFit(exprs_pair, design_1)

# <---- Model 3: exprs corrected for methylation ~ tissue
# specify design matrix
resid_exprs <- array(0, dim = dim(exprs_pair))
for (index in 1:nrow(exprs_pair)){
  resid_exprs[index,] <- lm(t(exprs_pair[index, ]) ~ t(methyl_pair[index, ]))$resid
}
rownames(resid_exprs) <- rownames(exprs_pair)
model_3 <- lmFit(resid_exprs, design_1)  


## Effect size differences

# get effect sizes
beta1 <- coef(model_1[,2])
beta3 <- coef(model_3[,2])
se_beta1 <- se_beta2 <- se_beta3 <- cov_beta13 <- vector("numeric", ngenes)
# <---- get variances of beta1^S tissue effect
se_beta1 <- model_1$sigma*sqrt((solve(t(design_1)%*%design_1))[2,2])
head(cbind(se_beta1,model_1$sigma*sqrt(model_1$cov.coefficients[2,2])))
# <---- get variances of beta2 methylation effect
for (g in 1:length(se_beta2)) {
  design_2g <- model.matrix(~ as.numeric(methyl_pair[g,]))  
  sigma_g <- model_1$sigma[g]
  se_beta2[index] <- sigma_g*sqrt((solve(t(design_2g)%*%design_2g))[2,2])
}
# <---- get variances of beta3 tissue effect
A <- solve(t(design_1)%*%design_1)%*%t(design_1)
contr.vector <- array(c(0,1), dim=c(2,1))
# compute beta3 by hand
for (g in 1:length(se_beta3)) {
  M_g <- t(methyl_pair[g,])
  design_2g <- model.matrix(~ as.numeric(M_g))  
  A_2g <- solve(t(design_2g)%*%design_2g)%*%t(design_2g)
  sigma_g <- model_1$sigma[g]
  var_beta2g <- (se_beta2^2)[g]
  var_part1 <- (se_beta1[g])^2
  var_part2 <- ( A%*%M_g%*%var_beta2g%*%t(M_g)%*%t(A) )[2,2]
  var_part3 <- ( 2*(sigma_g^2)*A%*%t(A_2g)%*%contr.vector%*%t(M_g)%*%t(A) )[2,2]
  se_beta3[g] <- sqrt(var_part1 + var_part2 + var_part3)
  }
# cov(beta1,beta3)
for (g in 1:length(cov_beta13)) {
  M_g <- t(methyl_pair[g,])
  design_2g <- model.matrix(~ as.numeric(M_g))  
  A_2g <- solve(t(design_2g)%*%design_2g)%*%t(design_2g)
  sigma_g <- model_1$sigma[g]
  var_part1 <- (se_beta1[g])^2
  cov_beta13[g] <- var_part1 - (sigma_g^2)*((A %*% t(A_2g) %*%contr.vector%*%t(M_g)%*%t(A))[2,2])
}
  
# cov(beta1-beta3)
cov_diff_sqrt <- sqrt(se_beta1^2 + se_beta3^2 - 2*cov_beta13)
beta_diff <- beta1-beta3

## Run adaptive shrinkage (with Matthew Stephens' package ashr)

fit_ash <- ash(as.vector(beta_diff), cov_diff_sqrt, mode=0)
names(fit_ash$result)

# lfsr is the minimum of PositiveProb and NegativeProb
# look at genes at various cut-off, say svalue < .01 etc. do they make sense?

ind_sig <- (fit_ash$result$svalue < FSR_level)
ind_sig[ind_sig == TRUE] <- "red"
ind_sig[ind_sig == FALSE] <- "black"

make_plot <- cbind(beta1-beta3, cov_diff_sqrt, ind_sig, fit_ash$result$lfdr)
colnames(make_plot) <- c("Diff", "SE", "Sign", "LFDR")
make_plot[,1] <- as.numeric(make_plot[,1])
make_plot[,2] <- as.numeric(make_plot[,2])
make_plot[,3] <- as.character(make_plot[,3])
make_plot[,4] <- as.numeric(make_plot[,4])

return(make_plot)
}
```

## Function that will perform the intertissue analysis- non-DE genes
```{r}

effect_size_hc_tissue_non_DE <- function(two_tissues, FDR_level, FSR_level){
## Determine DE genes (for tissue analysis)

# Factors that will go in the lm
chimp_human_heart <- two_tissues
index_sample <- chimp_human_heart
FDR_level <- FDR_level

# Get variables
species <- as.data.frame(samples[,2])
tissue <- as.data.frame(samples[,3])
RIN <- as.data.frame(samples[,4])

# Sort tissue samples
  tissue <- as.data.frame(samples[,3])
  tissue <- tissue[chimp_human_heart,]
  tissue_no_extra <- droplevels.factor(tissue)

# Run the linear model in limma
  design <- model.matrix(~ as.factor(tissue_no_extra) + RIN[chimp_human_heart,])
  fit_all <- lmFit(cpm.voom.cyclic[,chimp_human_heart], design)
  fit_all <- eBayes(fit_all)

# Get results
  HvC_Heart_fit_all = topTable(fit_all, coef=2, adjust="BH", number=Inf, sort.by="none")
  HvC_Heart_fit_all_5perc <-
HvC_Heart_fit_all[which(HvC_Heart_fit_all$adj.P.Val < FDR_level), ]

# Find which DE genes have methylation values associated with it
    human_chimp_heart <-  rownames(exprs) %in% HvC_Heart_fit_all_5perc$genes
  human_chimp_heart <- as.data.frame(human_chimp_heart)
  counts_genes_in <- cbind(exprs, human_chimp_heart)
  counts_genes_in_cutoff <- subset(counts_genes_in, human_chimp_heart ==
"FALSE")
  counts_genes_in_cutoff <- counts_genes_in_cutoff[,1:79]
  expression_values_only <- counts_genes_in_cutoff[,2:32]
  methylation_values_only <- counts_genes_in_cutoff[,49:79]
  dim(expression_values_only)
  dim(methylation_values_only)
  
# Rename is that we can use Joyce's code
exprs_subset <- expression_values_only
methyl_subset <- methylation_values_only
  
# check agreement between methylation data and expression data
all.equal(rownames(methyl_subset), rownames(exprs_subset))
all.equal(colnames(methyl_subset), colnames(exprs_subset))

# <---- sample data matching
samples_sub <- samples[samples$Sample_ID %in% colnames(exprs_subset),]


exprs_pair <- exprs_subset[,index_sample]
methyl_pair <- methyl_subset[,index_sample]
species <- droplevels.factor(samples$Species[index_sample])
tissue <- droplevels.factor(samples$Tissue[index_sample])
RIN <- samples$RIN[index_sample]


# Extract parameters
N <- ncol(exprs_pair)
ngenes <- nrow(methyl_pair)
# <---- Model 1: exprs ~ tissue
# specify tissue coding
design_1 <- model.matrix(~tissue + RIN)
design_1[design_1[,2]==0,2] <- -1
model_1 <- lmFit(exprs_pair, design_1)

# <---- Model 3: exprs corrected for methylation ~ tissue
# specify design matrix
resid_exprs <- array(0, dim = dim(exprs_pair))
for (index in 1:nrow(exprs_pair)){
  resid_exprs[index,] <- lm(t(exprs_pair[index, ]) ~ t(methyl_pair[index, ]))$resid
}
rownames(resid_exprs) <- rownames(exprs_pair)
model_3 <- lmFit(resid_exprs, design_1)  


## Effect size differences

# get effect sizes
beta1 <- coef(model_1[,2])
beta3 <- coef(model_3[,2])
se_beta1 <- se_beta2 <- se_beta3 <- cov_beta13 <- vector("numeric", ngenes)
# <---- get variances of beta1^S tissue effect
se_beta1 <- model_1$sigma*sqrt((solve(t(design_1)%*%design_1))[2,2])
head(cbind(se_beta1,model_1$sigma*sqrt(model_1$cov.coefficients[2,2])))
# <---- get variances of beta2 methylation effect
for (g in 1:length(se_beta2)) {
  design_2g <- model.matrix(~ as.numeric(methyl_pair[g,]))  
  sigma_g <- model_1$sigma[g]
  se_beta2[index] <- sigma_g*sqrt((solve(t(design_2g)%*%design_2g))[2,2])
}
# <---- get variances of beta3 tissue effect
A <- solve(t(design_1)%*%design_1)%*%t(design_1)
contr.vector <- array(c(0,1), dim=c(2,1))
# compute beta3 by hand
for (g in 1:length(se_beta3)) {
  M_g <- t(methyl_pair[g,])
  design_2g <- model.matrix(~ as.numeric(M_g))  
  A_2g <- solve(t(design_2g)%*%design_2g)%*%t(design_2g)
  sigma_g <- model_1$sigma[g]
  var_beta2g <- (se_beta2^2)[g]
  var_part1 <- (se_beta1[g])^2
  var_part2 <- ( A%*%M_g%*%var_beta2g%*%t(M_g)%*%t(A) )[2,2]
  var_part3 <- ( 2*(sigma_g^2)*A%*%t(A_2g)%*%contr.vector%*%t(M_g)%*%t(A) )[2,2]
  se_beta3[g] <- sqrt(var_part1 + var_part2 + var_part3)
  }
# cov(beta1,beta3)
for (g in 1:length(cov_beta13)) {
  M_g <- t(methyl_pair[g,])
  design_2g <- model.matrix(~ as.numeric(M_g))  
  A_2g <- solve(t(design_2g)%*%design_2g)%*%t(design_2g)
  sigma_g <- model_1$sigma[g]
  var_part1 <- (se_beta1[g])^2
  cov_beta13[g] <- var_part1 - (sigma_g^2)*((A %*% t(A_2g) %*%contr.vector%*%t(M_g)%*%t(A))[2,2])
}
  
# cov(beta1-beta3)
cov_diff_sqrt <- sqrt(se_beta1^2 + se_beta3^2 - 2*cov_beta13)
beta_diff <- beta1-beta3

## Run adaptive shrinkage (with Matthew Stephens' package ashr)

fit_ash <- ash(as.vector(beta_diff), cov_diff_sqrt, mode=0)
names(fit_ash$result)

# lfsr is the minimum of PositiveProb and NegativeProb
# look at genes at various cut-off, say svalue < .01 etc. do they make sense?

ind_sig <- (fit_ash$result$svalue < FSR_level)
ind_sig[ind_sig == TRUE] <- "red"
ind_sig[ind_sig == FALSE] <- "black"

make_plot <- cbind(beta1-beta3, cov_diff_sqrt, ind_sig, fit_ash$result$lfdr)
colnames(make_plot) <- c("Diff", "SE", "Sign", "LFDR")
make_plot[,1] <- as.numeric(make_plot[,1])
make_plot[,2] <- as.numeric(make_plot[,2])
make_plot[,3] <- as.character(make_plot[,3])
make_plot[,4] <- as.numeric(make_plot[,4])

return(make_plot)
}
```

### Results: human heart versus human kidney

```{r}
######### Human heart versus human kidney ###############

hk_DE_make_plot <- effect_size_hc_tissue_DE(c(17, 20, 21, 24, 25, 28, 29), 0.05, 0.05)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human heart vs human kidney)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(17, 20, 21, 24, 25, 28, 29), 0.05, 0.05)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in human heart vs human kidney)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])

######### Human heart versus human kidney ###############

hk_DE_make_plot <- effect_size_hc_tissue_DE(c(17, 20, 21, 24, 25, 28, 29), 0.05, 0.01)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human heart vs human kidney)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.01", "s >= 0.01"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(17, 20, 21, 24, 25, 28, 29), 0.05, 0.01)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in human heart vs human kidney)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.01", "s >= 0.01"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])

######### Human heart versus human kidney ###############

hk_DE_make_plot <- effect_size_hc_tissue_DE(c(17, 20, 21, 24, 25, 28, 29), 0.1, 0.1)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human heart vs human kidney)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(17, 20, 21, 24, 25, 28, 29), 0.1, 0.1)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in human heart vs human kidney)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.1", "s >= 0.1"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])

```

### Results: chimp heart versus chimp kidney

```{r}
######### Chimp heart versus human chimp kidney ###############

hk_DE_make_plot <- effect_size_hc_tissue_DE(c(1, 2, 5, 6, 9, 10, 13, 14), 0.05, 0.05)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in chimp heart vs chimp kidney)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(1, 2, 5, 6, 9, 10, 13, 14), 0.05, 0.05)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in chimp heart vs chimp kidney)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])

######### Chimp heart versus human chimp kidney ###############

hk_DE_make_plot <- effect_size_hc_tissue_DE(c(1, 2, 5, 6, 9, 10, 13, 14), 0.05, 0.01)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in chimp heart vs chimp kidney)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.01", "s >= 0.01"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(1, 2, 5, 6, 9, 10, 13, 14), 0.05, 0.01)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in chimp heart vs chimp kidney)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.01", "s >= 0.01"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])

######### Chimp heart versus human chimp kidney ###############

hk_DE_make_plot <- effect_size_hc_tissue_DE(c(1, 2, 5, 6, 9, 10, 13, 14), 0.1, 0.1)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in chimp heart vs chimp kidney)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.1", "s >= 0.1"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(1, 2, 5, 6, 9, 10, 13, 14), 0.1, 0.1)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in chimp heart vs chimp kidney)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.1", "s >= 0.1"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])

```

### Human heart kidney versus human kidney

```{r}
hk_DE_make_plot <- effect_size_hc_tissue_DE(c(18, 20, 22, 24, 26, 28, 30), 0.05, 0.05)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human heart vs human liver)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

a <- sum(hk_DE_make_plot[,3] == "red")
a
b <- length(hk_DE_make_plot[,3])
b

hk_make_plot <- effect_size_hc_tissue_non_DE(c(18, 20, 22, 24, 26, 28, 30), 0.05, 0.05)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in human heart vs human liver)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

c <- sum(hk_make_plot[,3] == "red")
c
d <- length(hk_make_plot[,3])
d
```

### Chimp heart versus chimp liver

```{r}
hk_DE_make_plot <- effect_size_hc_tissue_DE(c(1, 3, 5, 7, 9, 11, 13, 15), 0.05, 0.05)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in chimp heart vs chimp liver)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(1, 3, 5, 7, 9, 11, 13, 15), 0.05, 0.05)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in chimp heart vs chimp liver)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])
```


### Human heart versus human lung

```{r}
hk_DE_make_plot <- effect_size_hc_tissue_DE(c(19, 20, 23, 24, 27, 28, 31), 0.05, 0.05)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human heart vs human lung)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(19, 20, 23, 24, 27, 28, 31), 0.05, 0.05)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in human heart vs human lung)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])
```

### Chimp heart versus chimp lung

```{r}
hk_DE_make_plot <- effect_size_hc_tissue_DE(c(1, 4, 5, 8, 9, 12, 13, 16), 0.05, 0.05)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in chimp heart vs chimp lung)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(1, 4, 5, 8, 9, 12, 13, 16), 0.05, 0.05)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in chimp heart vs chimp lung)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])
```


### Human kidney versus human liver

```{r}
hk_DE_make_plot <- effect_size_hc_tissue_DE(c(17, 18, 21, 22, 25, 26, 29, 30), 0.05, 0.05)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human kidney vs human liver)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(17, 18, 21, 22, 25, 26, 29, 30), 0.05, 0.05)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in human kidney vs human liver)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])
```

### Chimp kidney versus chimp liver

```{r}
hk_DE_make_plot <- effect_size_hc_tissue_DE(c(2, 3, 6, 7, 10, 11, 14, 15), 0.05, 0.05)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in chimp kidney vs chimp liver)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(2, 3, 6, 7, 10, 11, 14, 15), 0.05, 0.05)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in chimp kidney vs chimp liver)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])
```


### Human kidney versus human lung

```{r}
hk_DE_make_plot <- effect_size_hc_tissue_DE(c(17, 19, 21, 23, 25, 27, 29, 31), 0.05, 0.05)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human kidney vs human lung)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(17, 19, 21, 23, 25, 27, 29, 31), 0.05, 0.05)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in human kidney vs human lung",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])
```

### Chimp kidney versus chimp lung

```{r}
hk_DE_make_plot <- effect_size_hc_tissue_DE(c(2, 4, 6, 8, 10, 12, 14, 16), 0.05, 0.05)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in chimp kidney vs chimp lung)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(2, 4, 6, 8, 10, 12, 14, 16), 0.05, 0.05)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in chimp kidney vs chimp lung)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])
```


### Human liver versus human lung

```{r}
hk_DE_make_plot <- effect_size_hc_tissue_DE(c(18, 19, 22, 23, 26, 27, 30, 31), 0.05, 0.05)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in human liver vs human lung)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(18, 19, 22, 23, 26, 27, 30, 31), 0.05, 0.05)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in human liver vs human lung",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])
```

### Chimp liver versus chimp lung

```{r}
hk_DE_make_plot <- effect_size_hc_tissue_DE(c(3, 4, 7, 8, 11, 12, 15, 16), 0.05, 0.05)

plot(x=hk_DE_make_plot[,1], y=hk_DE_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (DE genes in chimp liver vs chimp lung)",
     pch=16, cex=.6, col = hk_DE_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_DE_make_plot[,3] == "red")
length(hk_DE_make_plot[,3])

hk_make_plot <- effect_size_hc_tissue_non_DE(c(3, 4, 7, 8, 11, 12, 15, 16), 0.05, 0.05)

plot(x=hk_make_plot[,1], y=hk_make_plot[,2],
     ylab = "Standard error of the difference",
     xlab = "Difference in effect size from 2 lms", 
     main = "Difference of effect size (non-DE genes in chimp liver vs chimp lung)",
     pch=16, cex=.6, col = hk_make_plot[,3])
legend("topleft", legend=c("s < 0.05", "s >= 0.05"), col=c("red", "black"), pch=16:16, cex=0.8)

sum(hk_make_plot[,3] == "red")
length(hk_make_plot[,3])
```
