---
title: "Filtering_analysis"
author: "Lauren Blake"
date: "June 10, 2016"
output: word_document
---

### The goal of this file is to establish filtering criteria for multispecies RNA-seq data. This analysis is exploratory; our final scheme can be found in the beginning of the GC content normalization file.

### Introduction

There are the main considerations for filtering:
  
1. **What to filter (raw counts or CPM).** Our lab frequently uses CPM in human RNA-seq and multi-species RNA-seq data (e.g. Gallego Romero and Pavlovic et al. 2015). 

2. **What threshold(s) to filter** at e.g. how many samples, tissues, and/or species should have a present value in order to include the gene in the later analysis. For example, Julien had used the cut-off of log2(cpm > 1) in at least 4 of the samples for each gene. This is a potentially lax cutoff and may lead to us having lots of missing data when we perform the differential expression analysis. We have 12 tissue-species pairs. Therefore, we will evaluate the performance of 2/4, 3/4, and 4/4 samples in all 12 tissue-species pairs. 

*Note: After this preliminary exploration, we SNP calling (via GATK) that the sample Human 1 Heart is Human 1 but most likely a Human 1 Liver. Therefore, in future analyses, we will drop this sample.*

```{r}
# Load libraries 

library("gplots")
library("ggplot2")
library("RColorBrewer")
library("scales")
library("edgeR")
library("R.utils")
library("plyr")
library("limma")
library("statmod")
library("EDASeq")
source("~/Reg_Evo_Primates/ashlar-trial/analysis/functions.R")

# Load colors 

colors <- colorRampPalette(c(brewer.pal(9, "Blues")[1],brewer.pal(9, "Blues")[9]))(100)
pal <- c(brewer.pal(9, "Set1"), brewer.pal(8, "Set2"), brewer.pal(12, "Set3"))

#Load the data

  #Raw counts

counts_genes <- read.delim("~/Reg_Evo_Primates/ashlar-trial/data/counts_genes.txt")

summary(counts_genes)

  #Sample information

samples <- read.csv("~/Reg_Evo_Primates/ashlar-trial/data/Sample_info_RNAseq.csv")

labels <- paste(samples$Species, samples$Tissue, sep=" ")

dim(counts_genes)
```

### Patterns of log2(CPM) 

We are beginning with 30030 genes and 48 samples. Let's do some preliminary analysis on the log2(CPM) to determine an appropriate cutoff


```{r}
# log2(CPM) adjusted for library sizes

dge_original <- DGEList(counts=as.matrix(counts_genes), genes=rownames(counts_genes), group = as.character(t(labels)))
dge_original <- calcNormFactors(dge_original)

expr_cutoff <- -5.5

tmm_cpm <- cpm(dge_original, normalized.lib.sizes=TRUE, log=TRUE, prior.count = 0.25)
head(tmm_cpm)

summary(tmm_cpm)
hist(tmm_cpm, main = "log2(CPM) values in unfiltered data (n = 48 samples)", breaks = 100, ylim = c(0, 50000), xlab = "log2(CPM) values")
abline(v = expr_cutoff, col = "red", lwd = 3)

# Median log(CPM) adjusted for library sizes

cpm_log <- cpm(counts_genes, normalized.lib.sizes=TRUE, log = TRUE, prior.count = 0.25)
median_log2_cpm <- apply(cpm_log, 1, median)
hist(median_log2_cpm, breaks = 100, ylim = c(0, 1500))
abline(v = expr_cutoff, col = "red", lwd = 3)

```

These plots are very similar.

### Patterns of log2(CPM) for each species

```{r}
# Looking at chimpanzees
hist(tmm_cpm[,1], main = "log2(CPM) values in unfiltered data (chimpanzee heart sample 1)", xlab = "log2(CPM)", breaks = 100, ylim = c(0, 5000))
abline(v = expr_cutoff, col = "red", lwd = 3)


hist(tmm_cpm[,1:16], main = "log2(CPM) values in unfiltered chimpanzee samples", breaks = 100, xlab = "log2(CPM)", ylim = c(0, 20000))
abline(v = expr_cutoff, col = "red", lwd = 3)

# Looking at humans
hist(tmm_cpm[,17], main = "log2(CPM) values in unfiltered data (human heart sample 1)", xlab = "log2(CPM)", breaks = 100, ylim = c(0, 5000))
abline(v = expr_cutoff, col = "red", lwd = 3)

hist(tmm_cpm[,17:32], main = "log2(CPM) values in unfiltered data (all human samples)", xlab = "log2(CPM)", breaks = 100, ylim = c(0, 20000))
abline(v = expr_cutoff, col = "red", lwd = 3)

# Rhesus macaques

# Looking at chimpanzees
hist(tmm_cpm[,33], main = "log2(CPM) values in unfiltered rhesus macaque heart sample 1", xlab = "log2(CPM)", breaks = 100, ylim = c(0, 5000))
abline(v = expr_cutoff, col = "red", lwd = 3)


hist(tmm_cpm[,33:48], main = "log2(CPM) vallues in rhesus macaque samples", xlab = "log2(CPM)", breaks = 100, ylim = c(0, 20000))
abline(v = expr_cutoff, col = "red", lwd = 3)

```


### Filtering 1: log2(CPM) > 1 in at least 4 samples (filtering scheme proposed by Julien)

What happens when we require keeping only those that have log2(CPM) > 1 in at least 4 samples? 

```{r}
cpm <- cpm(dge_original, normalized.lib.sizes=TRUE, log=TRUE, prior.count=0.25)
cpm_filtered <- cpm[rowSums(cpm > 1) >= 4, ]
dim(cpm_filtered)
```

This leaves us with 15,487 genes. 

### Different log2(cpm) thresholds (1 to 48 samples required)

```{r}

#Find number of genes for each sample threshold
filtered_genes <- array(1:48, dim = (c(48,1)))

for (i in 1:48){
  cpm_filtered <- cpm[rowSums(cpm > 1) >= i, ]
  filter_genes <- as.data.frame(dim(cpm_filtered))
  filtered_genes[i, 1] <- filter_genes[1,]
  
  }

# Plot the results
plot(filtered_genes, pch = 16, xlab = "Number of samples required to have log2(cpm) > 1", ylab = "Number of genes", main = "Genes for each sample threshold")
```

There are `r filtered_genes[48,1]` genes when we require log2(cpm) > 1 in all 48 samples. 

### Filtering 2: Log2(CPM) > a particular threshold adjusted for library size in 4/4 samples, 3/4 samples, and 2/4 samples per tissue-species pair. 

```{r}
# Assign all 12 tissue-species pairs
chimp_hearts <- c(1, 5, 9, 13)
chimp_kidneys <- c(2,6,10,14)
chimp_livers <- c(3,7,11,15)
chimp_lungs <- c(4,8,12,16)

human_hearts <- c(17,21,25,29)
human_kidneys <- c(18,22,26,30)
human_livers <- c(19,23,27,31)
human_lungs <- c(20,24,28,32)

rhesus_hearts <- c(33,37,41,45)
rhesus_kidneys <- c(34,38,42,46)
rhesus_livers <- c(35,39,43,47)
rhesus_lungs <- c(36,40,44,48)

# Save the counts
                          # Raw counts = log2(counts_genes)
                          # CPM = tmm_cpm
counts_genes_filtered_3 <- tmm_cpm

# Put the number of samples that you want to have for every tissue-species pair (count/4 samples)

count = 2
threshold = -5.5

counts_genes_filtered_3A <- counts_genes_filtered_3[rowSums(counts_genes_filtered_3[, chimp_hearts] >= threshold) >= count, ] 

#dim(counts_genes_filtered_3A)

counts_genes_filtered_3B <- counts_genes_filtered_3A[rowSums(counts_genes_filtered_3A[, chimp_kidneys] >= threshold) >= count, ] 

dim(counts_genes_filtered_3B)

counts_genes_filtered_3C <- counts_genes_filtered_3B[rowSums(counts_genes_filtered_3B[, chimp_livers] >= threshold) >= count, ] 

dim(counts_genes_filtered_3C)

counts_genes_filtered_3D <- counts_genes_filtered_3C[rowSums(counts_genes_filtered_3C[, chimp_lungs] >= threshold) >= count, ] 

dim(counts_genes_filtered_3D)

counts_genes_filtered_3E <- counts_genes_filtered_3D[rowSums(counts_genes_filtered_3D[, human_hearts] >= threshold) >= count, ] 

dim(counts_genes_filtered_3E)

counts_genes_filtered_3F <- counts_genes_filtered_3E[rowSums(counts_genes_filtered_3E[, human_kidneys] >= threshold) >= count, ] 

dim(counts_genes_filtered_3F)

counts_genes_filtered_3G <- counts_genes_filtered_3F[rowSums(counts_genes_filtered_3F[, human_livers] >= threshold) >= count, ] 

dim(counts_genes_filtered_3G)

counts_genes_filtered_3H <- counts_genes_filtered_3G[rowSums(counts_genes_filtered_3G[, human_lungs] >= threshold) >= count, ] 

dim(counts_genes_filtered_3H)

counts_genes_filtered_3I <- counts_genes_filtered_3H[rowSums(counts_genes_filtered_3H[, rhesus_hearts] >= threshold) >= count, ] 

dim(counts_genes_filtered_3I)

counts_genes_filtered_3J <- counts_genes_filtered_3I[rowSums(counts_genes_filtered_3I[, rhesus_kidneys] >= threshold) >= count, ] 

dim(counts_genes_filtered_3J)

counts_genes_filtered_3K <- counts_genes_filtered_3J[rowSums(counts_genes_filtered_3J[, rhesus_livers] >= threshold) >= count, ] 

dim(counts_genes_filtered_3K)

counts_genes_filtered_3L <- counts_genes_filtered_3K[rowSums(counts_genes_filtered_3K[, rhesus_lungs] >= threshold) >= count, ] 

dim(counts_genes_filtered_3L)

#counts_genes_filtered_3L_sums <- rowSums(counts_genes_filtered_3L)
#hist(counts_genes_filtered_3L_sums, main = "Number of genes remaining (2/4 species-tissue criteria)", xlab = "Number of genes")
```

**When we require each sample in each tissue-species pair to have log2(CPM) > -5 in each gene, we have 14,090 genes remaining.**

**When we require 3/4 samples in each tissue-species pair to have log2(CPM) > -5 in each gene, we have 15,776 genes remaining.** 

**When we require 2/4 samples in each tissue-species pair to have log2(CPM) > -5 in each gene, we have 17,109 genes remaining.** 

For comparison,

When we require each sample in each tissue-species pair to have log2(CPM) > 0 in each gene, we have 9,803 genes remaining. 

When we require 3/4 samples in each tissue-species pair to have log2(CPM) > 0 in each gene, we have 10,579 genes remaining. 

When we require 2/4 samples in each tissue-species pair to have log2(CPM) > 0 in each gene, we have 10,157 genes remaining. 


### Distribution of NAs/0 counts

We can get an idea about the distribution of the missing data (where log2(CPM) > threshold) e.g. if one species or tissue-species pair is over-represented in the missing genes

```{r}
# Copy the raw counts table
counts_genes_filtered <- tmm_cpm

# If the raw count in a cell is > threshold, replace the counts with a 1
counts_genes_filtered[counts_genes_filtered > threshold ] <- 1

# If the raw count in a cell is =< threshold, replace the counts with a 0
counts_genes_filtered[counts_genes_filtered <= threshold ] <- 0
head(counts_genes_filtered)

# Count the number of genes that have a raw count < 1
x <- 30030 - colSums(counts_genes_filtered)

par(las=1)
plot(x, pch = 16, ylab = "Number of Genes with log2(CPM) < -5", xlab = "Sample name", xaxt = 'n', col = as.factor(samples$Tissue), main = "Number of genes with 0 raw counts/sample")
axis(side = 1, at = 1:48,labels = samples$Species)
legend("topright", c("Heart", "Kidney", "Liver", "Lung"), lwd = c(2,2,2,2), col = c("black", "red", "green", "blue"))
```


### Density plots of gene counts when we require 2/4 samples in each tissue-species pair to have log2(CPM) > -5.5

```{r}

# Find the original counts of all of the genes that fit the 2/4 criteria and then re-do TMM

inshared_lists = row.names(counts_genes) %in% rownames(counts_genes_filtered_3L)
inshared_lists_data <- as.data.frame(inshared_lists)
counts_genes_in <- cbind(counts_genes, inshared_lists_data)
counts_genes_in_2_of_4 <- subset(counts_genes_in, inshared_lists_data == "TRUE")
counts_genes_in_2_of_4 <- counts_genes_in_2_of_4[,1:48]

# Take the TMM of the genes that meet the criteria of 2/4 in 12 tissue-species pairs

dge_2_in_4 <- DGEList(counts=as.matrix(counts_genes_in_2_of_4), genes=rownames(counts_genes_in_2_of_4), group = as.character(t(labels)))
dge_2_in_4 <- calcNormFactors(dge_2_in_4)

cpm <- cpm(dge_2_in_4, normalized.lib.sizes=TRUE, log=TRUE, prior.count = 0.25)
head(cpm)

summary(cpm)
hist(cpm, xlab = "Log2(CPM)", main = "Log2(CPM) values for genes meeting the filtering criteria", breaks = 100 )
```

```{r}
all_hearts <- c(1,5,9,13,17, 21,25,29,33,37,41,45)
all_kidneys <- c(2,6,10,14,18, 22,26,30,34,38,42,46)
all_livers <- c(3,7,11,15,19, 23,27,31,35,39,43,47)
all_lungs <- c(4,8,12,16,20, 24,28,31,36,40,44,48)

col = as.data.frame(pal[as.numeric(samples$Species)])
col_hearts = col[all_hearts,]
col_kidneys = col[all_kidneys, ]
col_livers = col[all_livers, ]
col_lungs = col[all_lungs, ]

group = as.data.frame(samples$Species)
group_hearts = group[all_hearts, ]
group_kidneys = group[all_kidneys, ]
group_livers = group[all_livers, ]
group_lungs = group[all_lungs, ]

plotDensities(cpm[,all_hearts], col=col_hearts, legend = FALSE)
legend('topright', legend = levels(group_hearts), col = levels(col_hearts), pch = 20)

plotDensities(cpm[,all_kidneys], col=col_kidneys, legend = FALSE)
legend('topright', legend = levels(group_kidneys), col = levels(col_kidneys), pch = 20)

plotDensities(cpm[,all_livers], col=col_livers, legend = FALSE)
legend('topright', legend = levels(group_livers), col = levels(col_livers), pch = 20)

plotDensities(cpm[,all_lungs], col=col_lungs, legend = FALSE)
legend('topright', legend = levels(group_lungs), col = levels(col_lungs), pch = 20)

# Look at all tissues in each species

plotDensities(cpm[,1:16], group=samples$Tissue, col=pal[as.numeric(samples$Tissue)], legend = "topright")

plotDensities(cpm[,17:32], group=samples$Tissue, col=pal[as.numeric(samples$Tissue)], legend = "topright")

plotDensities(cpm[,33:48], group=samples$Tissue, col=pal[as.numeric(samples$Tissue)], legend = "topright")

```

